---
title: 'Section B: State of Washington School Finance Systems and Structures'
author: "Katherine Chang"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: html_document
        toc: true
        toc_float: true
        fig_width: 8
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache.lazy = FALSE,
                      fig.align = "center"
                      )
```

# Introduction

```{r data, include=FALSE}
library(tidyverse)
library(googledrive)
library(haven)

id_f196 <- drive_get(as_id("https://drive.google.com/file/d/1sGIIj3hplOHq7qPL7S7YTmQByzuCcHVU/view?usp=sharing"))

df_f196 <- id_f196 %>%
  drive_read_raw() %>%
  read_dta()

df_f196 %>% count(schyear)

df_f196_trim <- df_f196%>%
  select(
    "coudis",
    "schyear",
    "rev_stloc_18d_Capital",         
    "rev_stloc_18d_Debt",             
    "rev_stloc_18d_General",          
    "rev_stloc_18d_Transport",       
    "rev_stloc_18d_Total",            
    "rev_st_18d_Capital",             
    "rev_st_18d_Debt",               
    "rev_st_18d_General",             
    "rev_st_18d_Transport",           
    "rev_st_18d_Total",
    "staterev_staffsal_r19_2019_18d", 
    "staterev_cissal_r19_2019_18d",  
    "staterev_tchsal_r19_2019_18d",   
    "staterev_tchsalsm_r19_2019_18d", 
    "staterev_staffsal_r20_2019_18d",
    "staterev_cissal_r20_2019_18d",   
    "localrev_201819Spring",          
    "localrev_201819Spring_18d", 
    "dis_enroll",
    "cpi18")                          

```

Our guiding research questions include:

- How have Washington public schools been funded in the past few years (NEED YEARS), particularly for districts serving high-poverty and students of color? 
- How have McCleary school finance reforms changed schools funding sources per pupil? 
- How have these changes varied across districts? 
- How do schools spend their resources?


```{r revenue, include=FALSE}



```


```{r expenditures, include=FALSE}



```


```{r mccleary, include=FALSE}



```


```{r demographics, include=FALSE}



```


```{r maps, include=FALSE}

library(tidycensus)
library(tigris)

```

