---
title: 'Section A: Student-level Changes during COVID-19'
author: "Alex Liu"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_width: 8
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache.lazy = FALSE,
                      fig.align = "center"
                      )
```

#Notes
## 11/2 
### need to find income indicators for course taking
### for spatial plots: reason for missing data (grey areas): 1. missing and suppression at district level; 2. some districts have data for only one of the SYs, then the change is going to be NA; 3. some districts in the shapefile do not appear in the ospi stats 

# Introduction

```{r data, include=FALSE}
library(tidyverse)
library(googledrive)

id_enrollment <- drive_get(as_id("https://drive.google.com/file/d/1hfVC_BgnUF6lBJYcGf6lhtMQOxEdig2v/view?usp=sharing"))
id_graduation_pathways <- drive_get(as_id("https://drive.google.com/file/d/1hgO_VNQ2DdDtPBAa_Mtdkq1JhwPSD4uI/view?usp=sharing"))
id_reopening_survey <- drive_get(as_id("https://drive.google.com/file/d/1YbyBi1LfHz36aLyVItJs6mJ4nNWxuUkd/view?usp=sharing"))
id_course_outcomes_by_student <- drive_get(as_id("https://drive.google.com/file/d/1hfA_RgRAldfsLXzH6hg776kV_x96xDgI/view?usp=sharing"))
id_course_outcomes_by_outcome <- drive_get(as_id("https://drive.google.com/file/d/1hcUTvX5xYIVyejpyS73z0Xo-64w2uRCZ/view?usp=sharing"))
id_discipline <- drive_get(as_id("https://drive.google.com/file/d/1hkh1eV3QJ0ZkJRRB1VIlUIaJDE4yYoYF/view?usp=sharing"))

df_enrollment <- id_enrollment %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)
#https://data.wa.gov/Education/OSPI-Enrollment-Changes-2019-20-to-2020-21/5u4d-iirb

df_graduation_pathways <- id_graduation_pathways %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)
#https://data.wa.gov/Education/Data-Displays-Graduation-Pathways-OSPI/nxmy-bggx

df_reopening_survey <- id_reopening_survey %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)
#https://data.wa.gov/dataset/School-Facility-ReOpening-Survey/9i5d-c2m8

df_course_outcomes_by_student <- id_course_outcomes_by_student %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)
#https://data.wa.gov/Education/Student-Course-Outcomes-by-Student/f4u4-uezp

df_course_outcomes_by_outcome <- id_course_outcomes_by_outcome %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)
#https://data.wa.gov/dataset/Student-Course-Outcomes-by-Outcome/cjza-wqsm

df_discpline <- id_discipline %>%
  drive_read_string() %>%
  read_csv(col_name=TRUE)

#OSPI Suppression Rules
#https://k12.wa.us/sites/default/files/public/StudentInformation/SuppressionRulesforPublicReporting.pdf

```

```{r libraries, include=FALSE}
library(tidyverse)
library(ggplot2)
library(formattable)
library(reshape2)
library(grid)
library(gridExtra)
library(sf)
library(spData)

#library(rJava)
#library(tabulizer) 
```

# Enrollment
### Have trouble with grade-specific df, need to work on scrape from nested pdf (do-able but time consuming). once the datasets ready, the plot I &II codes can be used with minimal editing 10/26

## plot I: counts; year 2021; sorted by races; sorted by income levels; bar = month; grades = kindergaiden & 6th & 9th

```{r enrollment counts all students, include=FALSE}

month_in_df <- c("October", "November", "December", "January")

enrollment_by_races <- df_enrollment %>%
  filter(StudentGroupType == "Race" | StudentGroupType == "All") %>%
  filter(Measure == "2021 Enrolled") %>%
  select(!c(StudentGroupType, Measure)) 
  
long_enroll_by_races <- melt(enrollment_by_races, 
                             id.vars = "StudentGroup",
                             measure.vars = month_in_df,
                             variable.name = "Month",
                             value.name="EnrollmentCounts") %>%
  arrange(StudentGroup, Month) 

#enrollment_by_races <- as.data.frame(t(enrollment_by_races)) 
#names(enrollment_by_races) <- enrollment_by_races[1, ]
#enrollment_by_races <- enrollment_by_races[-1, ] 

#sapply(long_enroll_by_races, class)
#chars <- sapply(long_enroll_by_races, is.character)
#long_enroll_by_races[ , chars] <- as.data.frame(apply(long_enroll_by_races[ , chars], 2, as.numeric))

plot_enroll_count <- long_enroll_by_races %>%
  filter(StudentGroup != "All Students") %>%
  ggplot(aes(fill = Month, 
             x = StudentGroup, 
             y = EnrollmentCounts)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Enrollment Counts for SY2020-21") +
        labs(x = "Student Groups", 
             y = "Enrollment Counts") +
  theme(axis.text.y = element_text(angle = 45)) +
  theme_minimal() +
  coord_flip()


plot_enroll_count #need to reorder fill
```

```{r load enrollment data, include=FALSE}
id_enrollment_kids <- drive_get(as_id("https://drive.google.com/file/d/1lCa5ap7xqWXTslj7kEqIQfCLNX0WJf0q/view?usp=sharing"))
id_enrollment_six <- drive_get(as_id("https://drive.google.com/file/d/1KF-9IeKo7xrVJ3Ny5pUl8hKaax-expSE/view?usp=sharing"))
id_enrollment_nine <- drive_get(as_id("https://drive.google.com/file/d/1uysRPNVca71FFiTEb36cVqTzblTiwXmK/view?usp=sharing"))
id_change_enrollment_kids <- drive_get(as_id("https://drive.google.com/file/d/1oFO1A9FLJ_QPtT9IBEuiO1e6gzvRiHDI/view?usp=sharing"))
id_change_enrollment_six <- drive_get(as_id("https://drive.google.com/file/d/197DNjQPhvx4zS23C_SWInq9tw3nvoiXP/view?usp=sharing"))
id_change_enrollment_nine <- drive_get(as_id("https://drive.google.com/file/d/1SZkJZlYMDnwHtMIl0GfqhbIGXosoP9-U/view?usp=sharing"))

df_enrollment_kids <- id_enrollment_kids %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)

df_enrollment_six <- id_enrollment_six %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)

df_enrollment_nine <- id_enrollment_nine %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)

df_change_enrollment_kids <- id_change_enrollment_kids %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)

df_change_enrollment_six <- id_change_enrollment_six %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)

df_change_enrollment_nine <- id_change_enrollment_nine %>%
  drive_read_string() %>%
  read_csv(col_name=TRUE)
```


```{r enrollment counts kindergarden}

df_enrollment_kids[2,1] <- "All Students"
enroll_race_kids <- df_enrollment_kids %>%
  #put wrong labels when scraping
  select(c("X", "October_9th", 
           "November_9th", "December_9th", 
           "January_9th", "February_9th")) %>%
  filter(X %in% c("All Students",
               "American Indian/ Alaskan Native",
               "Asian",
               "Black/ African American",
               "Hispanic/ Latino of any race(s)",
               "Native Hawaian/ Other Pacific Islander",
               "Two or More Races",
               "White"))

names(enroll_race_kids) <- c("Student Group", "October", "November", 
                            "December", "January", "February")

months_in_df <- c("October", "November", "December", "January", "February")
long_enroll_race_kids <- melt(enroll_race_kids, 
                             id.vars = "Student Group",
                             measure.vars = months_in_df,
                             variable.name = "Month",
                             value.name=c("Enrollment")) %>%
  arrange(`Student Group`, Month) 

plot_enroll_count <- long_enroll_by_races %>%
  filter(StudentGroup != "All Students") %>%
  ggplot(aes(fill = Month, 
             x = StudentGroup, 
             y = EnrollmentCounts)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Enrollment Counts for SY2020-21") +
        labs(x = "Student Groups", 
             y = "Enrollment Counts") +
  theme(axis.text.y = element_text(angle = 45)) +
  theme_minimal() +
  coord_flip()
```


```{r enrollment counts sixth grade}
df_enrollment_six[2,2] <- "All Students"
enroll_race_six <- df_enrollment_six %>%
  select(c("X", "October_6th", 
           "November_6th", "December_6th", 
           "January_6th", "February_6th")) %>%
  filter(X %in% c("All Students",
               "American Indian/ Alaskan Native",
               "Asian",
               "Black/ African American",
               "Hispanic/ Latino of any race(s)",
               "Native Hawaian/ Other Pacific Islander",
               "Two or More Races",
               "White"))

names(enroll_race_six) <- c("Student Group", "October", "November", 
                            "December", "January", "February")

months_in_df <- c("October", "November", "December", "January", "February")
long_enroll_race_six <- melt(enroll_race_six, 
                             id.vars = "Student Group",
                             measure.vars = months_in_df,
                             variable.name = "Month",
                             value.name=c("Enrollment")) %>%
  arrange(`Student Group`, Month) 

plot_enroll_count <- long_enroll_by_races %>%
  filter(StudentGroup != "All Students") %>%
  ggplot(aes(fill = Month, 
             x = StudentGroup, 
             y = EnrollmentCounts)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Enrollment Counts for SY2020-21") +
        labs(x = "Student Groups", 
             y = "Enrollment Counts") +
  theme(axis.text.y = element_text(angle = 45)) +
  theme_minimal() +
  coord_flip()
```


```{r enrollment counts ninth grade}

df_enrollment_nine[2,1] <- "All Students"
enroll_race_nine <- df_enrollment_nine %>%
  select(c("X", "October_9th", 
           "November_9th", "December_9th", 
           "January_9th", "February_9th")) %>%
  filter(X %in% c("All Students",
               "American Indian/ Alaskan Native",
               "Asian",
               "Black/ African American",
               "Hispanic/ Latino of any race(s)",
               "Native Hawaian/ Other Pacific Islander",
               "Two or More Races",
               "White"))

names(enroll_race_nine) <- c("Student Group", "October", "November", 
                            "December", "January", "February")

months_in_df <- c("October", "November", "December", "January", "February")
long_enroll_race_nine <- melt(enroll_race_nine, 
                             id.vars = "Student Group",
                             measure.vars = months_in_df,
                             variable.name = "Month",
                             value.name=c("Enrollment")) %>%
  arrange(`Student Group`, Month) 
```


## plot II: try pie charts? datatype: % = each race/all student; year 2021; each monthly; comparison across different grades
```{r enrollment percentage all students, include=FALSE}
enrollment_perc_races <- data.frame(matrix(NA,    
                          nrow = nrow(enrollment_by_races),
                          ncol = ncol(enrollment_by_races)))

dimnames(enrollment_perc_races) <- list(rownames(enrollment_by_races), 
                                        colnames(enrollment_by_races))


for (j in 2:ncol(enrollment_by_races)) {
  for (i in nrow(enrollment_by_races):1) {
    enrollment_perc_races[i, j] = percent(enrollment_by_races[i, j]/enrollment_by_races[1,j]*100)
    enrollment_perc_races[i, 1] = enrollment_by_races[i, 1]
  }
}

long_enroll_perc_races <- melt(enrollment_perc_races, 
                             id.vars = "StudentGroup",
                             measure.vars = month_in_df,
                             variable.name = "Month",
                             value.name="EnrollmentPercentage") %>%
  arrange(StudentGroup, Month) 

plot_enroll_perc <- long_enroll_perc_races %>%
  filter(StudentGroup != "All Students") %>%
  ggplot(aes(fill = Month, x = StudentGroup, y = EnrollmentPercentage)) + 
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  coord_flip()

plot_enroll_perc

#add col "other" for pie charts

```

```{r enrollment percentage kindergarden}

all_students_kids <- long_enroll_race_kids %>%
  filter(`Student Group` == "All Students") %>%
  select("Enrollment")

long_enroll_race_kids <- long_enroll_race_kids %>%
  group_by(`Student Group`) %>%
  mutate("Percentage" = Enrollment/all_students_kids) %>%
  ungroup()  %>%
  arrange(`Student Group`, Month)

```

```{r enrollment percentage sixth grade}

all_students_six <- long_enroll_race_six %>%
  filter(`Student Group` == "All Students") %>%
  select("Enrollment")

long_enroll_race_six <- long_enroll_race_six %>%
  group_by(`Student Group`) %>%
  mutate("Percentage" = Enrollment/all_students_six) %>%
  ungroup()  %>%
  arrange(`Student Group`, Month)
```

```{r enrollment percentage ninth grade}

all_students_nine <- long_enroll_race_nine %>%
  filter(`Student Group` == "All Students") %>%
  select("Enrollment")

long_enroll_race_nine <- long_enroll_race_nine %>%
  group_by(`Student Group`) %>%
  mutate("Percentage" = Enrollment/all_students_nine) %>%
  ungroup()  %>%
  arrange(`Student Group`, Month)

```


## plot III: splitted bar w/ % changes; low-income vs. non-low-income; sorted by years or races?

```{r enrollment change kindergarden race}

change_enroll_kids_race <- df_change_enrollment_kids %>%
  filter(StudentGroup %in% c("All Students",
                             "American Indian/ Alaskan Native",
                             "Asian",
                             "Black/ African American",
                             "Hispanic/ Latino of any race(s)",
                             "Native Hawaian/ Other Pacific Islander",
                             "Two or More Races",
                             "White"))

months_in_df <- c("October", "November", "December", "January", "February")
long_change_enroll_kids_race <- melt(change_enroll_kids_race, 
                                     id.vars = "StudentGroup",
                                     measure.vars = months_in_df,
                                     variable.name = "Month",
                                     value.name=c("Change in Enrollment")) %>%
  arrange(`StudentGroup`, Month) 

```

```{r enrollment change kindergarden income}

change_enroll_kids_income <- df_change_enrollment_kids %>%
  filter(StudentGroup %in% c("All Students",
                             "Low-Income",
                             "Non-Low Income"))

months_in_df <- c("October", "November", "December", "January", "February")
long_change_enroll_kids_income <- melt(change_enroll_kids_income, 
                                     id.vars = "StudentGroup",
                                     measure.vars = months_in_df,
                                     variable.name = "Month",
                                     value.name=c("Change in Enrollment")) %>%
  arrange(`StudentGroup`, Month) 
```

```{r enrollment change sixth grade race}

change_enroll_six_race <- df_change_enrollment_six %>%
  filter(StudentGroup %in% c("All Students",
                             "American Indian/ Alaskan Native",
                             "Asian",
                             "Black/ African American",
                             "Hispanic/ Latino of any race(s)",
                             "Native Hawaian/ Other Pacific Islander",
                             "Two or More Races",
                             "White"))

months_in_df <- c("October", "November", "December", "January", "February")
long_change_enroll_six_race <- melt(change_enroll_six_race, 
                                     id.vars = "StudentGroup",
                                     measure.vars = months_in_df,
                                     variable.name = "Month",
                                     value.name=c("Change in Enrollment")) %>%
  arrange(`StudentGroup`, Month) 

```

```{r enrollment change sixth grade income}

change_enroll_six_income <- df_change_enrollment_six %>%
  filter(StudentGroup %in% c("All Students",
                             "Low-Income",
                             "Non-Low Income"))

months_in_df <- c("October", "November", "December", "January", "February")
long_change_enroll_six_income <- melt(change_enroll_six_income, 
                                     id.vars = "StudentGroup",
                                     measure.vars = months_in_df,
                                     variable.name = "Month",
                                     value.name=c("Change in Enrollment")) %>%
  arrange(`StudentGroup`, Month) 
```

```{r enrollment change ninth grade race}

change_enroll_nine_race <- df_change_enrollment_nine %>%
  filter(StudentGroup %in% c("All Students",
                             "American Indian/ Alaskan Native",
                             "Asian",
                             "Black/ African American",
                             "Hispanic/ Latino of any race(s)",
                             "Native Hawaian/ Other Pacific Islander",
                             "Two or More Races",
                             "White"))

months_in_df <- c("October", "November", "December", "January", "February")
long_change_enroll_nine_race <- melt(change_enroll_nine_race, 
                                     id.vars = "StudentGroup",
                                     measure.vars = months_in_df,
                                     variable.name = "Month",
                                     value.name=c("Change in Enrollment")) %>%
  arrange(`StudentGroup`, Month) 

```

```{r enrollment change ninth grade income}

change_enroll_nine_income <- df_change_enrollment_nine %>%
  filter(StudentGroup %in% c("All Students",
                             "Low-Income",
                             "Non-Low Income"))

months_in_df <- c("October", "November", "December", "January", "February")
long_change_enroll_nine_income <- melt(change_enroll_nine_income, 
                                     id.vars = "StudentGroup",
                                     measure.vars = months_in_df,
                                     variable.name = "Month",
                                     value.name=c("Change in Enrollment")) %>%
  arrange(`StudentGroup`, Month) 

```


# Absent
##[Students of color or low income students are further disadvantaged during the pandemic in terms of abs. Particularly lower grade levels ]
##Grade K, 3rd abs, 6th, 8th grade enrollment, 9th, 10th grade.

## plot I: absence count; year 2021 by races; grades: kindergargen, 3, 6, 8, 9, 10; 
```{r absence count kindergarden, include=FALSE}

```

```{r absence count 3rd grade, include=FALSE}

```

```{r absence count 6th grade, include=FALSE}

```

```{r absence count 8th grade, include=FALSE}

```

```{r absence count 9th grade, include=FALSE}

```

```{r absence count 10th grade, include=FALSE}

```

## plot II: absence change; datatype: % changed; 2019 vs 2021 (ignore 2020); grades: kindergargen, 3, 6, 8, 9, 10; sorted by races and income levels

```{r absence change kindergarden, include=FALSE}

```

```{r absence change 3rd grade, include=FALSE}

```

```{r absence change 6th grade, include=FALSE}

```

```{r absence change 8th grade, include=FALSE}

```

```{r absence change 9th grade, include=FALSE}

```

```{r absence change 10th grade, include=FALSE}

```

# Course Taking 
## plot I: counts; only use "non-passing grades"; sorted by income levels; 2020-21; bar; w/ and w/o reference; reference = state ave 

```{r course taking}
#code need to be generalized and plots need to be touched up

course_by_race <- df_course_outcomes_by_outcome %>%
  filter(`Organization Name` == "Washington State" & 
           `School Year` == "2020-21" & 
           !is.na(`All Non-Passing Grades`)) %>%
  select_if(~!all(is.na(.))) %>%
  mutate(`All Non-Passing Grades` =  percent(`All Non-Passing Grades`))

state_ave_20_21 <- percent('11.0%')
  
course_by_race$`Grade Level` <- 
  factor(course_by_race$`Grade Level`, levels = c("9", "10", "11", "12", "All Grades"))

#p <- list()

ggplot(course_by_race,
       aes(fill= `Grade Level`,
                          x= `Student Group`, 
                          y=`All Non-Passing Grades`)) + 
  geom_bar(position="dodge", stat="identity") +
  geom_hline(yintercept = state_ave_20_21) + 
  theme_minimal() +
  coord_flip()


course_by_race <- df_course_outcomes_by_outcome %>%
  filter(`Organization Name` == "Washington State" & 
           `School Year` == "2019-20" & 
           !is.na(`All Non-Passing Grades`)) %>%
  select_if(~!all(is.na(.))) %>%
  mutate(`All Non-Passing Grades` =  percent(`All Non-Passing Grades`))

state_ave_19_20 <- percent('4.8%')
  
course_by_race$`Grade Level` <- 
  factor(course_by_race$`Grade Level`, levels = c("9", "10", "11", "12", "All Grades"))

ggplot(course_by_race,
       aes(fill= `Grade Level`,
                          x= `Student Group`, 
                          y=`All Non-Passing Grades`)) + 
  geom_bar(position="dodge", stat="identity") +
  geom_hline(yintercept = state_ave_19_20) + 
  theme_minimal() +
  coord_flip()


course_by_race <- df_course_outcomes_by_outcome %>%
  filter(`Organization Name` == "Washington State" & 
           `School Year` == "2018-19" & 
           `Student Group` != "Asian" & 
           !is.na(`All Non-Passing Grades`)) %>%
  select_if(~!all(is.na(.))) %>%
  mutate(`All Non-Passing Grades` =  percent(`All Non-Passing Grades`))

state_ave_18_19 <- percent('7.4%')
  
course_by_race$`Grade Level` <- 
  factor(course_by_race$`Grade Level`, levels = c("9", "10", "11", "12", "All Grades"))

ggplot(course_by_race,
       aes(fill= `Grade Level`,
                          x= `Student Group`, 
                          y=`All Non-Passing Grades`)) + 
  geom_bar(position="dodge", stat="identity") +
  geom_hline(yintercept = state_ave_18_19) + 
  theme_minimal() +
  coord_flip()

#g_three_year <- arrangeGrob(p[[1]], p[[2]], p[[3]], nrow = 3)

#g_three_year
```


## plot II. (for all students) 
## i: Calculate the Difference in all non-passing grade from 2018-19 to 2020-21, separately for four groups: low-income, non-low-income, students of color (excluding Asian), and White). Then generate a map below. By this graph, we get which districts have the largest change and for which group, numerically. 

```{r 3.2 differences btwn sy in non-passing by district}
course_geo <- df_course_outcomes_by_outcome %>%
  select(c(`Organization Name`, `School Year`, 
           `Student Group`, `Grade Level`, 
           `All Non-Passing Grades`)) %>%
  filter(!is.na(`All Non-Passing Grades`) &
           `School Year` != "2019-20") %>% 
  mutate(`All Non-Passing Grades` =  percent(`All Non-Passing Grades`))

#cannot use dcast due to uneven length for each year:
#split the df by school year; rename `All Non-Passing Grades` column by SY; 
course_geo_1819 <- filter(course_geo, `School Year` != "2020-21") 
colnames(course_geo_1819)[5] <- "SY2018-19"
  
course_geo_2021 <- filter(course_geo, `School Year` == "2020-21")
colnames(course_geo_2021)[5] <- "SY2020-21"

#merge yearly dfs and calculate the change
course_geo_wide <- merge(x = course_geo_2021, y = course_geo_1819, 
                         by = c("Organization Name", "Student Group", "Grade Level"), 
                         all = TRUE) %>%
  select(!c("School Year.x", "School Year.y")) %>%
  mutate(Change = `SY2018-19` - `SY2020-21` ) %>%
  filter(!is.na(Change)) %>%
  #filter(Change >= -1) %>%
  mutate(Race = NA) 

for (i in 1:nrow(course_geo_wide)) {
  if (course_geo_wide[i, "Student Group"] == "White") {
    course_geo_wide[i, "Race"] <- "White"
  } else if (course_geo_wide[i, "Student Group"] == "Asian") {
    course_geo_wide[i, "Race"] <- "Asian"
  } else {
    course_geo_wide[i, "Race"] <- "Non-white"
  }
}

course_groupby_race <- aggregate(course_geo_wide$Change, 
                           list(course_geo_wide$`Organization Name`, 
                                course_geo_wide$`Grade Level`,
                                course_geo_wide$Race), 
                           FUN=mean)

names(course_groupby_race) <- c("Districts", "Grade Level", "Race", "Change")
course_groupby_race <- course_groupby_race %>%
  arrange(`Grade Level`, Districts, Race) %>%
  filter(Race != "Asian" & Districts != "Washington State") 

# Shape file
wa_district_shape <- st_read("/Users/alexliu/Documents/GitHub/evidenceforequity/A/enrollment_absence_dataset/Washington_School_District_Boundaries-shp")
#wa_district_coord <- st_read("/Users/alexliu/Documents/GitHub/evidenceforequity/A/enrollment_absence_dataset/Washington_State_Public_Schools")

all_students <- course_groupby_race %>%
  filter(`Grade Level` == "All Grades")
all_grades_white <- all_students %>%
  filter(Race == "White")
all_grades_nonwhite <- all_students %>%
  filter(Race == "Non-white")

course_white_sf <- left_join(wa_district_shape, all_grades_white, by = c("LEAName" = "Districts"))
course_nonwhite_sf <- left_join(wa_district_shape, all_grades_nonwhite, by = c("LEAName" = "Districts"))

#plot for all grade levels; white students
ggplot()+
  geom_sf(data = course_white_sf, aes(fill = Change))+
  theme_bw()

#plot for all grade levels; nonwhite students (exclude asians)
ggplot()+
  geom_sf(data = course_nonwhite_sf, aes(fill = Change))+
  theme_bw()




ggplot() + 
  geom_sf(data = all_grades_white, 
          aes(fill = Change), size = .2) +
  scale_fill_distiller(name = "change in non-passing grades", palette = "YlOrRd") + 
  theme_bw()

  theme(panel.background = element_blank(),
        legend.title = element_text(size = 10),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 10),
        plot.caption = element_text(hjust = 0, size = 7, face = "italic")) +
  labs(title="Chicago Public School Safety",
       subtitle = "School safety and culture climate ratings in zip codes",
       caption = "Safety ratings range from 1 to 5 from very weak to very strong. 
Dots on map indicate schools that have not organized school climate.
Grey areas indicate lack of data.")


```


```{r 3.2 match district names with shape file, include=FALSE}
#matching district names
df_name <- unique(course_groupby_race$Districts)
shape_name <- sort(unique(wa_district_shape$LEAName))
`%!in%` <- Negate(`%in%`)
for (i in 1:length(df_name)) {
  if (df_name[i] %!in% shape_name) {
    print(i)
    print(df_name[i])
  }
}
for (j in 1:length(shape_name)) {
  if (shape_name[j] %!in% df_name) {
    print(j)
    print(shape_name[j])
  }
}
#district names need to be(can be) manually adjusted for geometry data: 
#df_name[108]"North Beach School District No. 64" = shape_name[159] "North Beach School District"
#df_name[151]"Seattle School District No. 1" = shape_name[220]"Seattle Public Schools"
#df_name[187]"Wellpinit School District #49" = shape_name[279]"Wellpinit School District"
## 7 schools in our dataset have no matches in shape file


```


## plot II (for all students)
### ii Calculate the gap in all non-passing in 2020-21 between low-income and non-low income (students of color (excluding Asian) and white) for each district and graph the difference using the map below.

```{r 3.3 differences btwn student types in non-passing by district}

course_compare <- df_course_outcomes_by_outcome %>%
  select(c(`Organization Name`, `School Year`, 
           `Student Group`, `Grade Level`, 
           `All Non-Passing Grades`)) %>%
  mutate(`All Non-Passing Grades` =  percent(`All Non-Passing Grades`)) %>%
  filter(!is.na(`All Non-Passing Grades`) &
          `School Year` == "2020-21" &
          `Student Group` != "Asian") 

for (i in 1:nrow(course_compare)) {
  if (course_compare[i, "Student Group"] == "White") {
    course_compare[i, "Race"] <- "White"
  } else {
    course_compare[i, "Race"] <- "Non-white"
  }
}

course_compare_groupby_race <- aggregate(course_compare$`All Non-Passing Grades`, 
                           list(course_compare$`Organization Name`, 
                                course_compare$`Grade Level`,
                                course_compare$Race), 
                           FUN=mean)
names(course_compare_groupby_race) <- c("Districts", 
                                        "Grade Level", 
                                        "Race", 
                                        "Ave All Non-Passing Grades")

course_compare_white <- filter(course_compare_groupby_race, Race == "White") 
colnames(course_compare_white)[4] <- "Ave All Non-Passing Grades White"
course_compare_nonwhite <- filter(course_compare_groupby_race, Race == "Non-white") 
colnames(course_compare_nonwhite)[4] <- "Ave All Non-Passing Grades Non_White"

course_diff_by_race<- merge(x = course_compare_white, 
                            y = course_compare_nonwhite, 
                         by = c("Districts", "Grade Level"), 
                         all = TRUE) %>%
  select(!c("Race.x", "Race.y")) %>%
  mutate(Change = `Ave All Non-Passing Grades White` - `Ave All Non-Passing Grades Non_White`) %>%
  filter(!is.na(Change)) %>%
  #filter(Change >= -1) %>%
  arrange(`Grade Level`, Districts) %>%
  filter(Districts != "Washington State") 

# Shape file
wa_district_shape <- st_read("/Users/alexliu/Documents/GitHub/evidenceforequity/A/enrollment_absence_dataset/Washington_School_District_Boundaries-shp")
#wa_district_coord <- st_read("/Users/alexliu/Documents/GitHub/evidenceforequity/A/enrollment_absence_dataset/Washington_State_Public_Schools")

diff_all_students <- course_diff_by_race %>%
  filter(`Grade Level` == "All Grades")

course_race_diff_sf <- left_join(wa_district_shape, diff_all_students, by = c("LEAName" = "Districts"))

ggplot()+
  geom_sf(data = course_race_diff_sf, aes(fill = Change))+
  theme_bw()

```


## plot III: 
* 9th grade on track: did not get a grade of D or F. The % of students that received grades of A,B, C.
* Regular attendance, meaning that has more than 90% attendance during the year
* Dual credit (not prestigious): (AP/IB courses is more prestigious); therefore, less useful
* COVID-19 modified (beginning of the school year to march, 2019-20 school year). The “modified” may indicate that. then OSPI created a fall to march dataset for multiple years of data

```{r course_outcomes, include=FALSE}

df_course_outcomes_by_student

df_course_outcomes_by_outcome
```

## plot IV; Graduation (the class of 2020). 
https://www.k12.wa.us/data-reporting/data-displays/graduation-pathways-display
Different pathways by income and students of color 
```{r graduation, include=FALSE}

df_graduation_pathways

```

