---
title: 'Section A-D: District and Union Leader Survey Analysis'
author: "Katherine Chang"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_width: 10
knit: (function(input, encoding) rmarkdown::render(input, output_file = "survey_analysis.html"))
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache.lazy = FALSE,
                      fig.align = "center"
                      )

library(here)
source(here("theme_uw.R"))
palette_uw <- colorRampPalette(c("#4B2E83", "#E8E3D3"))
palette_uw_reverse<- colorRampPalette(c("#E8E3D3", "#4B2E83"))
palette_black_white <- colorRampPalette(c("#FFFFFF", "#000000"))
```

# Introduction

```{r data_read, include=FALSE}

library(tidyverse)
library(googledrive)
library(haven)
library(reshape2)
library(readxl)


id_survey_district <- drive_get(as_id("https://drive.google.com/file/d/1iYMdwSU1TlVbtXCJu83DweK4rl65R6sc/view?usp=sharing"))

id_survey_union <- drive_get(as_id("https://drive.google.com/file/d/1TWXuCzc3UA3cGSkphqqZUyN9d1ovblp4/view?usp=sharing")) 

df_survey_district <- id_survey_district %>%
  drive_read_raw() %>%
  read_csv()

df_survey_union <- id_survey_union %>%
  drive_read_raw() %>%
  read_csv()

survey_district_questions = t(df_survey_district[1,])
survey_union_questions = t(df_survey_union[1,])

```

```{r demographics, include=FALSE}

district <- df_survey_district%>%
  slice(3:n())%>%
  select(-c("StartDate", "EndDate", "Status", "IPAddress", "Progress",
            "Duration (in seconds)", "Finished", "RecordedDate", "ResponseId",
            "RecipientLastName", "RecipientFirstName", "RecipientEmail", "LocationLatitude",
            "LocationLongitude", "DistributionChannel", "UserLanguage"))

district_melt <- melt(district, id.vars = "ExternalReference")

district_melt%>%
  filter(variable=="Q1")%>%
  group_by(value)%>%
  mutate(value_count=n())%>%
  ggplot(aes(x=reorder(value,-value_count))) +
  geom_bar(stat="count") +
  labs(title="District Survey Respondents", subtitle="Current and Primary Role", x="") +
  scale_x_discrete(labels = c("Superintendent", "Finance-related Admin", "Other", "NA"))+
  theme_uw()

district_melt%>%
  filter(variable=="Q2")%>%
  group_by(value)%>%
  mutate(value_count=n())%>%
  ggplot(aes(x=reorder(value,-value_count))) +
  geom_bar(stat="count") +
  labs(title="District Survey Respondents", subtitle="Years of Experience", x="") +
  theme_uw()

```

```{r finance, include=FALSE}

district_melt%>%
  filter(grepl("^Q3_[0-9]{1,2}$", variable))%>%
  na.omit(variable)%>%
  group_by(value)%>%
  mutate(value_count=n())%>%
  ggplot(aes(x=reorder(value,-value_count))) +
  geom_bar(stat="count") +
  labs(title="Top three most influential decision-makers at the district central office\nduring planned the 2020-21 school year budget planning",
       x="") +
  scale_x_discrete(labels = c("Superintendent", 
                              "Asst Superintendent/Chief of Staff", 
                              "School Board Members", 
                              "Superintendent's Cabinet Members", 
                              "School Principals",
                              "Budget Advisory Committee",
                              "Other",
                              "Union Leadership",
                              "Teachers",
                              "External Consultants",
                              "Community Members"))+
  theme_uw() +
  theme(axis.text.x=element_text(angle=15))

district_melt%>%
  filter(grepl("^Q4_[0-9]{1,2}$", variable))%>%
  na.omit(variable)%>%
  group_by(value)%>%
  mutate(value_count=n())%>%
  ggplot(aes(x=reorder(value,-value_count))) +
  geom_bar(stat="count") +
  labs(title="Top 3 sources of information  used by the district central office\n to develop the 2020-21 school year budget?", x="") +
  theme_uw() +
  theme(axis.text.x=element_text(angle=5))

district_melt%>%
  filter(grepl("^Q5_[0-9]{1,2}$", variable))%>%
  na.omit(variable)%>%
  group_by(value)%>%
  mutate(value_count=n())%>%
  ggplot(aes(x=reorder(value,-value_count))) +
  geom_bar(stat="count") +
  #labs(title="Top 3 sources of information  used by the district central office\n to develop the 2020-21 school year budget?", x="") +
  theme_uw() +
  theme(axis.text.x=element_text(angle=15))


district_melt%>%
  filter(grepl("^Q26_[0-9]{1,2}$", variable))%>%
  na.omit(variable)%>%
  group_by(value)%>%
  mutate(value_count=n())%>%
  ggplot(aes(x=reorder(value,-value_count))) +
  geom_bar(stat="count") +
  labs(title="Role districts had difficulty filling for the 2020-21 school year", x="") +
  theme_uw() +
  theme(axis.text.x=element_text(angle=15))


```

```{r union, include=FALSE}


```

Start with summary table of demographics

Frequency 
Mean
Standard Deviation

Correlate scatterplot, look at relationship between two variables

# Demographic of Survey Response
## Response
Roles
Years of Experience
Regional Representation 
Gender
Race

## District
Attendance Size
Student Racial Demographics
FRPL

Urbanicity, Regional differences (puget sound), by ESD

Look at the two surveys separately, and then link
