---
title: 'Section A-D: District and Union Leader Survey Analysis'
author: "Katherine Chang"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_width: 10
knit: (function(input, encoding) rmarkdown::render(input, output_file = "survey_analysis.html"))
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache.lazy = FALSE,
                      fig.align = "center"
                      )
theme_efe <-
    ggplot2::theme_classic() +
    ggplot2::theme(
      )
```

# Introduction

```{r data_read, include=FALSE}

library(tidyverse)
library(googledrive)
library(haven)


id_survey_district <- drive_get(as_id("https://drive.google.com/file/d/1XabcDxFGcUSF28KvT1BRGY_79N9Vyog9/view?usp=sharing"))

id_survey_union <- drive_get(as_id("https://drive.google.com/file/d/1TWXuCzc3UA3cGSkphqqZUyN9d1ovblp4/view?usp=sharing")) 

df_survey_district <- id_survey_district %>%
  drive_read_raw() %>%
  read_csv()

df_survey_union <- id_survey_union %>%
  drive_read_raw() %>%
  read_csv()

survey_district_questions = t(df_survey_district[1,])
survey_union_questions = t(df_survey_union[1,])

```

Start with summary table of demographics

Frequency 
Mean
Standard Deviation

Correlate scatterplot, look at relationship between two variables

# Demographic of Survey Response
## Response
Roles
Years of Experience
Regional Representation 
Gender
Race

## District
Attendance Size
Student Racial Demographics
FRPL

Urbanicity, Regional differences (puget sound), by ESD

Look at the two surveys separately, and then link

# Section B

Q3
Q3_11_TEXT
When your district central office planned the 2020-21 school year budget, who were the top three most influential decision-makers? - Selected Choice

Q4
What information was used by the district central office to develop the 2020-21 school year budget? Select the top three most important sources of information.

Q5
What factors represented notable challenges when your district central office planned the 2020-21 school year budget?

Q28
How did your district allocate federal stimulus funds (i.e., CARES, CRRSAA, ARP)?

Q27
How has your 2020-21 school year budget addressed students' learning disparities by income or race/ethnicity?

# Section C

Q26
Q26_13_TEXT
Which of the following roles did you have difficulty filling for the 2020-21 school year? Select all that apply. - Selected Choice

-Q15_1, Q15_10
To what extent do you agree with the following statements about the union for certificated teachers in your district? 

Q9
In the past five years, have you been part of the collective bargaining negotiation process for certificated teachers in your district?

Q10
Q10_11_TEXT
Since 2018, what were the district’s priorities during the most recent major collective bargaining for certificated teachers (major bargaining period not related to COVID-19 school reopening)?  Select the top three priorities from your district’s perspective. - Selected Choice

## COVID
Q18
Has your district engaged in negotiation with the union for school reopening during the Spring 2021 term?

Q12
What were your district’s priorities during union negotiations for Spring 2021 reopening? Select the top five priorities. - Selected Choice

Q12_14_TEXT
What were your district’s priorities during union negotiations for Spring 2021 reopening? Select the top five priorities. - Other(s), please specify - Text

Q13
Has your district engaged in negotiation with the union for school reopening for the Fall 2021 term?

Q14
What are your district’s priorities during union negotiations for Fall 2021 reopening? Select the top five priorities. - Selected Choice

Q14_14_TEXT
What are your district’s priorities during union negotiations for Fall 2021 reopening? Select the top five priorities. - 


## Section D