---
title: 'Section D: Impact and Implications of COVID-specific Policies/Funds'
author: "Alex Liu"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    fig_width: 8
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment = NA, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache.lazy = FALSE,
                      fig.align = "center"
                      )
```


```{r data, include=FALSE}
library(tidyverse)
library(googledrive)

id_LEA_plans <- drive_get(as_id("https://drive.google.com/file/d/1Yc35em3KL-06Cau-Iy0GBdvyPLO01CRu/view?usp=sharing"))


df_LEA_plans <- id_LEA_plans %>%
  drive_read_string() %>%
  read_csv(col_names=TRUE)
```

```{r libraries, include=FALSE}
library(googledrive)
library(tidyverse)
library(ggplot2)
library(formattable)
library(reshape2)
library(grid)
library(gridExtra)
library(sf)
library(spData)
library(leaidr)
library(stringr)
`%!in%` <- Negate(`%in%`)
#library(rJava)
#library(tabulizer) 
```

```{r uw_palette, include=FALSE}
# from purple to gold (low=gold, high=purple)
uw_palette <- colorRampPalette(c("#4b2e83", "#e8e3d3"))
uw_reverse <- colorRampPalette(c("#e8e3d3", "#4b2e83"))
black_white <- colorRampPalette(c("white", "black"))
```

* Note 12/14
  + ~~change percent to count~~
  + D1 ~~fix labels~~
  + D2 
  + D3 ~~explore student groups used by districts~~
  + D4 ~~generate figures~~

# D.1 EquityAnalysisToolName

> Summarize the frequency, bar chart, the categories of EquityAnalysisToolName, by unit of districts

* 305 school districts with 271 equity analysis tool names
  + many names are similar. are they the same?
  + see equityana.xlsx & equitytoolkeywords.csv

```{r D.1 EquityAnalysisToolName top}
df_equityana <- df_LEA_plans %>%
  select(c("LEAName", "EquityAnalysisToolName"))

df_equityana_unique <- unique(df_equityana[c("LEAName", "EquityAnalysisToolName")])

num_district <- nrow(df_equityana_unique)

#df_equityana_unique$LEAName[duplicated(df_equityana_unique$LEAName)]

#df_equityana_sep <- as.data.frame(str_split_fixed(df_equityana_unique$EquityAnalysisToolName, 
#                                                 ",",
#                                                 max(unlist(lapply(strsplit(df_equityana_unique$EquityAnalysisToolName, 
#                                                                            ","), 
#                                                                   length)))))
#names(df_equityana_sep) <- paste0("tool_", 1:ncol(df_equityana_sep))
#df_equityana_sep$LEAName <- df_equityana_unique$LEAName
#df_equityana_sep$tool_1[duplicated(df_equityana_sep$tool_1)]

df_equityana_insight <- df_equityana_unique %>% 
  filter(grepl("Insight", EquityAnalysisToolName, fixed = TRUE) |
         grepl("insight", EquityAnalysisToolName, fixed = TRUE)) %>%
  mutate(EquityAnalysisToolName = "Equity Rubric: Attending to Needs of the Most Vulnerable Populations Developed by Insight Education Group")

df_equityana_shoreline <- df_equityana_unique %>% 
  filter(grepl("Shoreline", EquityAnalysisToolName, fixed = TRUE) |
         grepl("shoreline", EquityAnalysisToolName, fixed = TRUE)) %>%
  mutate(EquityAnalysisToolName = "Shoreline Race & Equity Impact Decision-Making Tool")
  
df_equityana_count <- df_equityana_unique %>%
  mutate(EquityAnalysisToolName = case_when(grepl("insight", EquityAnalysisToolName, fixed = TRUE) ~ "Equity Rubric: Attending to Needs of the Most Vulnerable Populations Developed by Insight Education Group",
                                            grepl("Insight", EquityAnalysisToolName, fixed = TRUE) ~ "Equity Rubric: Attending to Needs of the Most Vulnerable Populations Developed by Insight Education Group",
                                            grepl("Attending to Needs", EquityAnalysisToolName, fixed = TRUE) ~ "Equity Rubric: Attending to Needs of the Most Vulnerable Populations Developed by Insight Education Group",
                                            grepl("Attending yo Needs", EquityAnalysisToolName, fixed = TRUE) ~ "Equity Rubric: Attending to Needs of the Most Vulnerable Populations Developed by Insight Education Group",
                                            grepl("Shoreline", EquityAnalysisToolName, fixed = TRUE) ~ "Shoreline Race & Equity Impact Decision-Making Tool",
                                            grepl("shoreline", EquityAnalysisToolName, fixed = TRUE) ~ "Shoreline Race & Equity Impact Decision-Making Tool",
                                            grepl("Equity Questions for Consideration", EquityAnalysisToolName, fixed = TRUE) ~ "Equity Questions for Consideration and Race & Equity Impact Decision-Making Tool",
                                            grepl("please note that we are in the process", EquityAnalysisToolName, fixed = TRUE) ~ "Hanover Research Toolkit (in process)",
                                            EquityAnalysisToolName == "7 Steps Pathway to Equity" ~ "7-Step Equity Tool",
                                            TRUE ~ EquityAnalysisToolName)) %>%
  
  group_by(EquityAnalysisToolName) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) 

#for (tools in df_equityana_count$EquityAnalysisToolName) {
#  print(grepl("Insight", tools, fixed = TRUE))
#}

df_equityana_count %>%
  filter(number >= 2) %>%
  ggplot(aes(x= reorder(EquityAnalysisToolName, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Equity Anlysis Tools") +
  labs(x = "", 
       y = "",
       caption = "Note: this figure includes equity analysis tools that are employed by more than one school district;
       305 school districts are included") +
  #scale_y_continuous(labels = percent, limits = c(0, 0.025)) +
  scale_x_discrete(labels = function(x) str_wrap(str_replace_all(x, "foo" , " "),
                                                 width = 70)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

```

```{r D.1 EquityAnalysisToolName modify, eval=FALSE, include=FALSE}
df_equityana_sep <- as.data.frame(str_split_fixed(df_equityana_unique$EquityAnalysisToolName, 
                                                 " ", max(unlist(lapply(strsplit(df_equityana_unique$EquityAnalysisToolName," "), length)))))
names(df_equityana_sep) <- paste0("word_", 1:ncol(df_equityana_sep))

df_equitytool_keyword <- df_equityana_sep %>% 
  gather(word, keyword) %>%
  replace(. == "", NA) %>%
  filter(!is.na(keyword)) %>%
  select(-word) %>%
  group_by(keyword) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) %>%
  filter(number >= 3) %>%
  filter(keyword %!in% c("the", "of", "and", "by", "&", 
                         "to", "-", "in", "is", "this",
                         "with", "using", "on", "An", "an",
                         "for", "from", "a", "The", "that",
                         "School", "Schools", "we", "We",
                         "our", "Lake", "Equity", "Tool", 
                         "District", "Tool,", "tool", "Valley",
                         "SD", "Analysis", "analysis",
                         "Decision-Making"))

df_equityana_modify_count <- mutate(df_equityana_count, keywords = "")

for (word in df_equitytool_keyword$keyword) {
  for (i in 1:nrow(df_equityana_modify_count)) {
    if (grepl(word, df_equityana_modify_count$EquityAnalysisToolName[i], fixed = TRUE)) {
    df_equityana_modify_count$keywords[i] <- paste(df_equityana_modify_count$keywords[i], word)
    }
  }
}

#write_csv(df_equityana_modify_count, "/Users/alexliu/Desktop/equitytoolkeywords.csv")


```

```{r D.1 EquityAnalysisToolName all, eval=FALSE, include=FALSE}

narrative_tools <- c("(1) intention with Technology for Learning (2) Family &amp; Community Partnerships (3) Program Evaluation/Audit of Assessment Systems to inform potential updates and ensure a comprehensive supportive Whole Child approach aligned with our OSD Student Outcomes and (4) updated system supports for continuous improvement practices for School Improvement Planning alongside District Improvement Planning and related monitoring for success ",
                     "Here in the Kahlotus School District we have been in an in-person traditional schedule since August 31 2020.  Due to our tiny student population we have been able to implement a traditional calendar and schedule with our students.  We practice daily interventions through Student and Teacher Access Time (STAT) for our students in grades 8-12.  These inerventions are designed and delivered by our certificated secondary teachers.  We also practice and deliver daily instruction through specific and targeted intervention to individual and small groups of students in grades k-6.  Our instruction in elementary classrooms is a natural environment where teaching is delivered at the targeted individual level. We only have 6 students in our k-1 classroom 7 students in our 2-3 classroom and 8 students in our 4-6 classroom.",
                     "Relationship with local health and human services for direct support around mental health issues and pandemic response/mitigation",
                     "Sources of Strength student programming mentor/partnership program for students of color additional special education and LTEL learner support Data engineer to improve data access and usability",
                     "The District initiated a new remote learning liaison position focused on students and families that choose the remote learning option.  This was a full time certificated staff member who was available to and for parents and students alike.  ",
                     "This plan is in early stages of development.  As we continue to get feedback from staff parents and community- and as we learn more about student needs- we will modify our approach for all students and for targeted student groups..",
                     "Wa He Lut Indian School is a small district that practices family tailored educational processes that reaches out to students and families utilizing school liasions")

df_equityana_count %>%
  filter(EquityAnalysisToolName %!in% narrative_tools) %>%
  ggplot(aes(x= reorder(EquityAnalysisToolName, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Equity Anlysis Tools") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included
       narrative equity analysis tools are not included in the figure: 
       1. (1) intention with Technology for Learning (2) Family & Community Partnerships (3) Program Evaluation/Audit of Assessment Systems to inform potential updates and ensure a comprehensive supportive Whole Child approach aligned with our OSD Student Outcomes and (4) updated system supports for continuous improvement practices for School Improvement Planning alongside District Improvement Planning and related monitoring for success.
       2. Here in the Kahlotus School District we have been in an in-person traditional schedule since August 31 2020.  Due to our tiny student population we have been able to implement a traditional calendar and schedule with our students.  We practice daily interventions through Student and Teacher Access Time (STAT) for our students in grades 8-12.  These inerventions are designed and delivered by our certificated secondary teachers.  We also practice and deliver daily instruction through specific and targeted intervention to individual and small groups of students in grades k-6.  Our instruction in elementary classrooms is a natural environment where teaching is delivered at the targeted individual level. We only have 6 students in our k-1 classroom 7 students in our 2-3 classroom and 8 students in our 4-6 classroom.
       3. Relationship with local health and human services for direct support around mental health issues and pandemic response/mitigation
       4. Sources of Strength student programming mentor/partnership program for students of color additional special education and LTEL learner support Data engineer to improve data access and usability.
       5. The District initiated a new remote learning liaison position focused on students and families that choose the remote learning option.  This was a full time certificated staff member who was available to and for parents and students alike.
       6. This plan is in early stages of development.  As we continue to get feedback from staff parents and community- and as we learn more about student needs- we will modify our approach for all students and for targeted student groups.
       7. Wa He Lut Indian School is a small district that practices family tailored educational processes that reaches out to students and families utilizing school liasions") +
  #scale_y_continuous(labels = percent, limits = c(0, 0.025)) +
  scale_x_discrete(labels = function(x) str_wrap(str_replace_all(x, "foo" , " "),
                                                 width = 70)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

```


# D.2 UniversalSupports

> Summarize the frequency of the categories of “UniversalSupports” (Bar chart), by unit of districts

* delete "Other - Write In (Required)"?
* "AVID" = "AVID 6-12"? "College in the High School + AVID"? 
* separate narratives?
    + "Here in the Kahlotus School District we have been in an in-person traditional schedule since August 31 2020.  Due
    to our tiny student population we have been able to implement a traditional calendar and schedule with our students.
    We practice daily interventions through Student and Teacher Access Time (STAT) for our students in grades 8-12. 
    These inerventions are designed and delivered by our certificated secondary teachers.  We also practice and deliver
    daily instruction through specific and targeted intervention to individual and small groups of students in grades
    k-6.  Our instruction in elementary classrooms is a natural environment where teaching is delivered at the targeted
    individual level. We only have 6 students in our k-1 classroom 7 students in our 2-3 classroom and 8 students in our
    4-6 classroom."
* "Summer learning resources for families and credit retrieval for high school students" = "Summer School"?
* see univsupp.xlsx

```{r D.2 UniversalSupports TOP26}

df_universalsupports <- df_LEA_plans %>%
  select(c("LEAName", "UniversalSupports")) 

df_univsupp_unique <- unique(df_universalsupports[c("LEAName", "UniversalSupports")])

#check whether one sd different inputs of plans
#df_universalsupports$LEAName[duplicated(df_universalsupprts$LEAName)]

df_univsupp_sep <- as.data.frame(str_split_fixed(df_univsupp_unique$UniversalSupports, 
                                                 ",", 
                                                 max(unlist(lapply(strsplit(df_univsupp_unique$UniversalSupports, ","), 
                                                                   length)))))

names(df_univsupp_sep) <- paste0("support_", 1:ncol(df_univsupp_sep))
df_univsupp_sep$LEAName <- df_univsupp_unique$LEAName

df_univsupp_count <- df_univsupp_sep %>% 
  gather(support, UniversalSupports, -LEAName) %>%
  replace(. == "", NA) %>%
  filter(!is.na(UniversalSupports)) %>%
  select(-support) %>%
  mutate(UniversalSupports = case_when(UniversalSupports == "CharacterStrong" ~ "Character Strong",
                                       UniversalSupports == "AVID " ~ "AVID",
                                       UniversalSupports == "District made Differentiated Interventions in Small Group" ~ "District Made Differentiated Interventions in Small Groups",
                                       UniversalSupports == "On line learning" ~ "Online learning",
                                       UniversalSupports == "Transition Supports ()" ~ "Transition Supports",
                                       UniversalSupports == "Strategic Staffing ()" ~ "Strategic Staffing",
                                       TRUE ~ UniversalSupports)) %>%
#  filter(UniversalSupports != "Other - Write In (Required)") %>%  #delelte Other?
  group_by(UniversalSupports) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) 

df_univsupp_count %>%
  slice(1:26) %>%
  ggplot(aes(x= reorder(UniversalSupports, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of the most employed Universal Supports") +
  labs(x = "", 
       y = "",
       caption = "Note: this figure includes all the plans that are employed by more than one school district;
       305 school districts are included") +
  #scale_y_continuous(labels = percent, limits = c(0, 1)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))


```

```{r D.2 UniversalSupports all}
narrative_plans <- c("Sources of Strength student programming mentor/partnership program for students of color additional special education and LTEL learner support Data engineer to improve data access and usability",
                     "The District initiated a new remote learning liaison position focused on students and families that choose the remote learning option.  This was a full time certificated staff member who was available to and for parents and students alike.  ",
                     "This plan is in early stages of development.  As we continue to get feedback from staff parents and community- and as we learn more about student needs- we will modify our approach for all students and for targeted student groups..",
                     "Social-emotional Learning/College Career Readiness daily meetings Second Step SEL curriculum K-5  structured early release",
                     "Relationship with local health and human services for direct support around mental health issues and pandemic response/mitigation",
                     "Here in the Kahlotus School District we have been in an in-person traditional schedule since August 31 2020.  Due to our tiny student population we have been able to implement a traditional calendar and schedule with our students.  We practice daily interventions through Student and Teacher Access Time (STAT) for our students in grades 8-12.  These inerventions are designed and delivered by our certificated secondary teachers.  We also practice and deliver daily instruction through specific and targeted intervention to individual and small groups of students in grades k-6.  Our instruction in elementary classrooms is a natural environment where teaching is delivered at the targeted individual level. We only have 6 students in our k-1 classroom 7 students in our 2-3 classroom and 8 students in our 4-6 classroom.",
                     "Wa He Lut Indian School is a small district that practices family tailored educational processes that reaches out to students and families utilizing school liasions",
                     "(1) intention with Technology for Learning (2) Family &amp; Community Partnerships (3) Program Evaluation/Audit of Assessment Systems to inform potential updates and ensure a comprehensive supportive Whole Child approach aligned with our OSD Student Outcomes and (4) updated system supports for continuous improvement practices for School Improvement Planning alongside District Improvement Planning and related monitoring for success ")

df_univsupp_count %>%
  filter(UniversalSupports %!in% narrative_plans) %>%
  ggplot(aes(x= reorder(UniversalSupports, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Universal Supports") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
#  scale_y_continuous(labels = percent, limits = c(0, 1)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
#  geom_text(aes(label = `% Completed` ), position = position_dodge(0.9),
#            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(hjust = 1, vjust =0.5, size=4.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))
```

# D.3 StrategicSupport
> Summarize the frequency of the categories of “StrategicSupport”,  by types of identified  student groups, by unit of district (we may not see a big difference across these types of students)

All the student groups in df:

StudentGroup                               | 
-------------------------------------------|---------------------------------------------------------
"American Indian/ Alaska Native students"  | "English Language Learners"                                            "Hispanic/Latino of any race(s) students"  | "Low-income students"                                                  "Students Experiencing Homelessness"       | "Students in Foster Care"                                              "Students with Disabilities"               | "Native Hawaiian/ Other Pacific Islander students"                     "Black/ African American students"         | "Asian students"                                                       "Two or more races students"               | "White students"                                                       
"African American Male Students"           | "All our students need some social-emotional support"
"Highly Capable"                           | "Students Furthest from Educational Justice"  
"n/a"                                      | "Students who are credit deficient"
"All students have been impacted."         | "Chronic or Extreme Chronic Absence"
"Gifted Students"                          | "High School AP & Running Start Students" 
"Students referred by teacher"             | "Students new to UPSD"
"Students below grade level performance targets" | "Our student group is the 43 students in K-5.  We have some students requiring intervention and are providing those services."
"With a total enrollment of 9 at the end of the 2020-21 school year all students were considered individually and as a group" |
                                                                                          
 
```{r D.3 StrategicSupport}

df_strategicsupport <- df_LEA_plans %>%
  select("LEAName", "StudentGroup", "StrategicSupports")

df_ss_unique <- unique(df_strategicsupport[c("LEAName", "StudentGroup", "StrategicSupports")]) %>%
  mutate(StudentGroup = case_when(StudentGroup == "Two or more races" ~ "Two or more races students",
                                  StudentGroup == "Students of two or More Races" ~ "Two or more races students",
                                  TRUE ~ StudentGroup))

fed_ethnicity <- c("American Indian/ Alaska Native students",
                   "Asian students",
                   "Black/ African American students",
                   "Hispanic/Latino of any race(s) students",
                   "Native Hawaiian/ Other Pacific Islander students",
                   "White students",
                   "Two or more races students",
                   "Low-income students",
                   "Students with Disabilities",
                   "Students Experiencing Homelessness",
                   "Students in Foster Care",
                   "English Language Learners")

df_ss_race <- filter(df_ss_unique, StudentGroup %in% fed_ethnicity)
df_ss_sep <- as.data.frame(str_split_fixed(df_ss_race$StrategicSupports,
                                           ",", 
                                           max(unlist(lapply(strsplit(df_ss_race$StrategicSupports, ","),
                                                             length)))))

num_support <- ncol(df_ss_sep)
names(df_ss_sep) <- paste0("support_", 1:num_support)
df_ss_sep$LEAName <- df_ss_race$LEAName 
df_ss_sep$StudentGroup <- df_ss_race$StudentGroup

for (i in 1:num_support){
  df_ss_sep[[i]] <- str_squish(df_ss_sep[[i]])
}

output <- vector("list", length(fed_ethnicity))
district_num <- vector("list", length(fed_ethnicity))

for (i in 1:length(fed_ethnicity)) {
  output[[i]] <- df_ss_sep %>%
    filter(StudentGroup == fed_ethnicity[i]) 
  
  district_num[[i]] <- as.numeric(nrow(output[[i]]))
  
  output[[i]] <- output[[i]] %>%
    gather(support, StrategicSupports, -LEAName) %>%
    replace(. == "", NA) %>%
    filter(!is.na(StrategicSupports)) %>%
    select(-support) %>%
    mutate(StrategicSupports = case_when(StrategicSupports == "Strategic Staffing (teacher advocates" ~ "Strategic Staffing (teacher advocates, advisory, looping)",
                                         StrategicSupports == "advisory" ~ "",
                                         StrategicSupports == "looping)" ~ "",
                                         TRUE ~ StrategicSupports)) %>%
    filter(StrategicSupports != "") %>%
    group_by(StrategicSupports) %>%
    summarise(number = n()) %>%
    arrange(desc(number)) %>%
    mutate(StudentGroup = fed_ethnicity[i]) %>%
    filter(StrategicSupports %!in% fed_ethnicity)
  
}

df_ss_count <- bind_rows(output)

df_ss_count %>%
  filter(StudentGroup %in% fed_ethnicity[1:3]) %>%
  ggplot(aes(x= reorder(StrategicSupports, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  facet_wrap(~StudentGroup, ncol = 3, labeller = label_wrap_gen(width=20)) +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Strategic Supports") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
#  scale_y_continuous(labels = numeric, limits = c(0, 1000)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
#  geom_text(aes(label = numeric), position = position_dodge(0.9),
#            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))

df_ss_count %>%
  filter(StudentGroup %in% fed_ethnicity[4:6]) %>%
  ggplot(aes(x= reorder(StrategicSupports, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  facet_wrap(~StudentGroup, ncol = 3, labeller = label_wrap_gen(width=20)) +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Strategic Supports") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
#  scale_y_continuous(labels = numeric, limits = c(0, 1000)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
#  geom_text(aes(label = `% Completed` ), position = position_dodge(0.9),
#            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))

df_ss_count %>%
  filter(StudentGroup %in% fed_ethnicity[7:9]) %>%
  ggplot(aes(x= reorder(StrategicSupports, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  facet_wrap(~StudentGroup, ncol = 3, labeller = label_wrap_gen(width=20)) +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Strategic Supports") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
#  scale_y_continuous(labels = numeric, limits = c(0, 1000)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
#  geom_text(aes(label = `% Completed` ), position = position_dodge(0.9),
#            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))

df_ss_count %>%
  filter(StudentGroup %in% fed_ethnicity[10:12]) %>%
  ggplot(aes(x= reorder(StrategicSupports, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  facet_wrap(~StudentGroup, ncol = 3, labeller = label_wrap_gen(width=20)) +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of Strategic Supports") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
#  scale_y_continuous(labels = numeric, limits = c(0, 1000)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
#  geom_text(aes(label = `% Completed` ), position = position_dodge(0.9),
#            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))

```

> Common student groups identified by school districts

```{r D.3 StrategicSupport student groups by grade levels, message=FALSE, warning=FALSE}

df_ss_stugroups <- df_LEA_plans %>%
  select("LEAName", "Grade", "StudentGroup")

df_ss_stugroups <- unique(df_ss_stugroups[c("LEAName", "Grade", "StudentGroup")]) %>%
  mutate(StudentGroup = case_when(StudentGroup == "Two or more races" ~ "Two or more races students",
                                  StudentGroup == "Students of two or More Races" ~ "Two or more races students",
                                  TRUE ~ StudentGroup)) 

grade_levels <- factor(c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "K"), 
                       levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "K"))

output <- vector("list", length(grade_levels))
for (i in seq_along(grade_levels)) {
  output[[i]] <- df_ss_stugroups %>%
    filter(Grade == grade_levels[i]) %>%
    group_by(StudentGroup, Grade) %>%
    summarise(number = n()) %>%
    arrange(desc(number))
}

long_group_name <- c("With a total enrollment of 9 at the end of the 2020-21 school year all students were considered individually and as a group",
                     "Our student group is the 43 students in K-5.  We have some students requiring intervention and are providing those services.")

ss_stugroups_count <- bind_rows(output) %>%
  filter(StudentGroup %!in% c("n/a", "NULL"))

ss_stugroups_count %>%
  filter(StudentGroup %!in% long_group_name) %>%
  filter(number > 1)%>%
  ggplot(aes(x= reorder(StudentGroup, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  facet_wrap(~as.factor(Grade), ncol=4, labeller = label_wrap_gen(width=20)) +
  ggtitle("Common student groups identified by school districts",
          subtitle = "sorted by grade levels") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
 # scale_y_continuous(labels = numeric, limits = c(0, 305)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
  geom_text(aes(label = number ), position = position_dodge(0.9),
            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))
```

```{r D.3 StrategicSupport student groups district counts}

ss_stugroup_district_count <- df_ss_stugroups %>%
    group_by(StudentGroup) %>%
    summarise(number = n()) %>%
    arrange(desc(number))


ss_stugroup_district_count %>%
  filter(StudentGroup %!in% c("n/a", "NULL"))%>%
  ggplot(aes(x= reorder(StudentGroup, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("Common student groups identified by school districts",
          subtitle = "for all grade levels") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
 # scale_y_continuous(labels = numeric, limits = c(0, 305)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
  geom_text(aes(label = number), position = position_dodge(0.9),
            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))
```


> Special student groups only used by single school district

School Districts use special student group |* marks schools that do not use derailed race/ethnicity groups
-------------------------------------------|------------------------------------------------------------------------
"Damman School District"*                  | Our student group is the 43 students in K-5.  We have some students requiring intervention and are providing those services. 
"Index Elementary School District 63"*     | All our students need some social-emotional support
"Seattle School District No. 1"            | African American Male Students; Highly Capable; Students Furthest from Educational Justice
"Shaw Island School District"*             | With a total enrollment of 9 at the end of the 2020-21 school year all students were considered individually and as a group
"Sultan School District"                   | All students have been impacted.
"Sumner School District"                   | Students who are credit deficient
"Tahoma School District"                   | Chronic or Extreme Chronic Absence
"Tukwila School District"                  | Gifted Students; High School AP & Running Start Students; Students referred by teacher
"University Place School District"         | Students below grade level performance targets; Students new to UPSD

```{r D.3 StrategicSupport special student groups, message=FALSE, warning=FALSE}

ss_stugroups_count %>%
  filter(StudentGroup %!in% long_group_name) %>%
  filter(number == 1)%>%
  ggplot(aes(x= reorder(StudentGroup, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  facet_wrap(~as.factor(Grade), ncol=4, labeller = label_wrap_gen(width=20)) +
  ggtitle("Student groups identified by single school districts",
          subtitle = "sorted by grade levels") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included;
       Two school districts identified their student group as (1) With a total enrollment of 9 at the end of the 2020-21 school year 
       all students were considered individually and as a group and (2) Our student group is the 43 students in K-5.
       We have some students requiring intervention and are providing those services.") +
 # scale_y_continuous(labels = numeric, limits = c(0, 305)) +
#  scale_fill_manual(name ="Universal Support", values = uw_reverse(26)) +
  geom_text(aes(label = number ), position = position_dodge(0.9),
            hjust = -0.1, vjust = 0.4, colour = "black", size = 1.5) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 5),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 9, colour = "black"))

spec_group <- ss_stugroups_count %>%
  filter(number == 1) 

spec_group_dist <- df_ss_stugroups %>%
  filter(StudentGroup %in% spec_group$StudentGroup)

```

# D.4 StrategiesForMentoring & Support

## StrategiesForMentoring

> Summarize the frequency of the “strategiesformentoring”, 

```{r D.4.i strategiesformentoring}

df_strategiesformentoring <- df_LEA_plans %>%
  select(c("LEAName", "StrategiesForMentoring")) 

df_sfm_unique <- unique(df_strategiesformentoring[c("LEAName", "StrategiesForMentoring")])

#check whether one sd different inputs of plans
#df_sfm_unique$LEAName[duplicated(df_sfm_unique$LEAName)]

df_sfm_sep <- as.data.frame(str_split_fixed(df_sfm_unique$StrategiesForMentoring, 
                                            ",", 
                                            max(unlist(lapply(strsplit(df_sfm_unique$StrategiesForMentoring, ","), 
                                                              length)))))

names(df_sfm_sep) <- paste0("mentor_", 1:ncol(df_sfm_sep))
df_sfm_sep$LEAName <- df_sfm_unique$LEAName

df_sfm_count <- df_sfm_sep %>%
  mutate(mentor_1 = str_squish(mentor_1),
         mentor_2 = str_squish(mentor_2),
         mentor_3 = str_squish(mentor_3)) %>%
  gather(mentor, StrategiesForMentoring, -LEAName) %>%
  replace(. == "", NA) %>%
  filter(!is.na(StrategiesForMentoring)) %>%
  select(-mentor) %>%
  mutate(StrategiesForMentoring = case_when(StrategiesForMentoring == "Strategic Staffing ()" ~ "Strategic Staffing",
                                          StrategiesForMentoring == "collaboration with stakeholders to target and combine wrap-around services with academic instruction for most-at risk students" ~ "*",
                                          TRUE ~ StrategiesForMentoring)) %>%
  group_by(StrategiesForMentoring) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) 

df_sfm_count %>%
  ggplot(aes(x= reorder(StrategiesForMentoring, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of strategies for mentoring") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included
       * collaboration with stakeholders to target and combine wrap-around
       services with academic instruction for most-at risk students") +
  #scale_y_continuous(labels = percent, limits = c(0, 1)) +
  #scale_x_discrete(labels = function(x) str_wrap(str_replace_all(x, "foo" , " "),
  #                                               width = 100)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(hjust=1, vjust =0.5, size=5.5),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

#ggsave("/Users/alexliu/Desktop/sfm.png")

```

## StrategiesForSupport

> The frequency of the “strategies for support”

```{r D.4.i strategiesforsupport}
df_strategiesforsupport <- df_LEA_plans %>%
  select(c("LEAName", "StrategiesForSupport")) 

df_sfs_unique <- unique(df_strategiesforsupport[c("LEAName", "StrategiesForSupport")]) 

#check whether one sd different inputs of plans
#df_sfs_unique$LEAName[duplicated(df_sfs_unique$LEAName)]

df_sfs_sep <- as.data.frame(str_split_fixed(df_sfs_unique$StrategiesForSupport, 
                                            ",", 
                                            max(unlist(lapply(strsplit(df_sfs_unique$StrategiesForSupport, ","), 
                                                              length)))))

names(df_sfs_sep) <- paste0("support_", 1:ncol(df_sfs_sep))
df_sfs_sep$LEAName <- df_sfs_unique$LEAName

df_sfs_count <- df_sfs_sep %>% 
  mutate(support_1 = str_squish(support_1),
         support_2 = str_squish(support_2),
         support_3 = str_squish(support_3)) %>%
  gather(support, StrategiesForSupport, -LEAName) %>%
  replace(. == "", NA) %>%
  filter(!is.na(StrategiesForSupport)) %>%
  select(-support) %>%
  mutate(StrategiesForSupport = case_when(StrategiesForSupport == "STAR/IAB's" ~ "Star/IAB's",
                                          StrategiesForSupport == "flixibility for TBIP to be used in multilingual and dual language programs to facilitate ELD development" ~ "*",
                                          StrategiesForSupport == "State support to find methods for supporting/retaining/and growing teachers of color through stipends and PD opportunities" ~ "**",
                                TRUE ~ StrategiesForSupport)) %>%
  group_by(StrategiesForSupport) %>%
  summarise(number = n()) %>%
  arrange(desc(number))

df_sfs_count %>%
  ggplot(aes(x= reorder(StrategiesForSupport, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "frequency of strategies for support") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included
       * flixibility for TBIP to be used in multilingual and dual
       language programs to facilitate ELD development
       ** State support to find methods for supporting/retaining/and
       growing teachers of color through stipends and PD opportunities") +
  #scale_y_continuous(labels = percent, limits = c(0, 1)) +
  #scale_x_discrete(labels = function(x) str_wrap(str_replace_all(x, "foo" , " "),
  #                                               width = 70), 
  #                 guide = guide_axis(check.overlap = TRUE))+
   #                guide = guide_axis(n.dodge=2)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(hjust=1, vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

#ggsave("/Users/alexliu/Desktop/sfs.png")
```

# Add-on: EngagingVoices

```{r Engaging all kinds of voices, echo=FALSE}

df_voice <- df_LEA_plans %>%
  select(c("LEAName", "EngagingStudentVoices", 
           "EngagingFamilyVoices", "EngagingCommunityVoices"))


df_voice_unique <- unique(df_voice[c("LEAName", 
                                     "EngagingStudentVoices",
                                     "EngagingFamilyVoices", 
                                     "EngagingCommunityVoices")])

#check whether one sd different inputs of plans
#df_universalsupports$LEAName[duplicated(df_universalsupprts$LEAName)]

df_voice_stu_unique <- df_voice_unique %>%
  select(c("LEAName", "EngagingStudentVoices")) 

df_voice_stu_sep <- as.data.frame(str_split_fixed(df_voice_stu_unique$EngagingStudentVoices, 
                                ",",
                                max(unlist(lapply(strsplit(df_voice_stu_unique$EngagingStudentVoices, ","),
                                                  length)))))

names(df_voice_stu_sep) <- paste0("student_", 1:ncol(df_voice_stu_sep))
df_voice_stu_sep$LEAName <- df_voice_stu_unique$LEAName

df_voice_fam_unique <- df_voice_unique %>%
  select(c("LEAName", "EngagingFamilyVoices"))

df_voice_fam_sep <- as.data.frame(str_split_fixed(df_voice_fam_unique$EngagingFamilyVoices, 
                                ",",
                                max(unlist(lapply(strsplit(df_voice_fam_unique$EngagingFamilyVoices, ","),
                                                  length)))))

names(df_voice_fam_sep) <- paste0("family_", 1:ncol(df_voice_fam_sep))
df_voice_fam_sep$LEAName <- df_voice_fam_unique$LEAName

df_voice_com_unique <- df_voice_unique %>%
  select(c("LEAName", "EngagingCommunityVoices"))

df_voice_com_sep <- as.data.frame(str_split_fixed(df_voice_com_unique$EngagingCommunityVoices, 
                                ",",
                                max(unlist(lapply(strsplit(df_voice_com_unique$EngagingCommunityVoices, ","),
                                                  length)))))

names(df_voice_com_sep) <- paste0("community_", 1:ncol(df_voice_com_sep))
df_voice_com_sep$LEAName <- df_voice_com_unique$LEAName

df_voice_stu_count <- df_voice_stu_sep %>% 
  gather(student, EngagingStudentVoices, -LEAName) %>%
  replace(. == "", NA) %>%
  filter(!is.na(EngagingStudentVoices)) %>%
  select(-student) %>%
  mutate(EngagingStudentVoices = str_squish(EngagingStudentVoices)) %>%
  group_by(EngagingStudentVoices) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) 

df_voice_fam_count <- df_voice_fam_sep %>% 
  gather(family, EngagingFamilyVoices, -LEAName) %>%
  replace(. == "", NA) %>%
  filter(!is.na(EngagingFamilyVoices)) %>%
  select(-family) %>%
  mutate(EngagingFamilyVoices = str_squish(EngagingFamilyVoices)) %>%
  group_by(EngagingFamilyVoices) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) 

df_voice_com_count <- df_voice_com_sep %>% 
    gather(community, EngagingCommunityVoices, -LEAName) %>%
  replace(. == "", NA) %>%
  filter(!is.na(EngagingCommunityVoices)) %>%
  select(-community) %>%
  mutate(EngagingCommunityVoices = str_squish(EngagingCommunityVoices)) %>%
  group_by(EngagingCommunityVoices) %>%
  summarise(number = n()) %>%
  arrange(desc(number)) 

df_voice_stu_count %>%
  ggplot(aes(x= reorder(EngagingStudentVoices, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "methods for engaging student voices") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
  #scale_y_continuous(labels = percent, limits = c(0, 1)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

df_voice_fam_count %>%
  ggplot(aes(x= reorder(EngagingFamilyVoices, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "methods for engaging family voices") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
  #scale_y_continuous(labels = percent, limits = c(0, 1)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

df_voice_com_count %>%
  ggplot(aes(x= reorder(EngagingCommunityVoices, number), 
             y= number)) + 
  geom_col(position="dodge", width = 0.7, fill="#4b2e83") +
  ggtitle("LEA Academic and Students Well-being Plans",
          subtitle = "methods for engaging community voices") +
  labs(x = "", 
       y = "",
       caption = "Note: 305 school districts are included") +
  #scale_y_continuous(labels = percent, limits = c(0, 1)) +
  theme_light() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.caption = element_text(hjust = 1, size = 9),
        legend.background = element_rect(fill = "white", size = 1, colour = "white"),
        legend.title=element_text(size = 10),
        legend.text=element_text(size = 7),
        axis.text.x  = element_text(angle=45, vjust=0.5, size=7),
        axis.text.y = element_text(vjust =0.5, size=7),
        axis.title.x  = element_text(size=7),
        axis.title.y  = element_text(size=12),
        strip.text.x = element_text(size = 12, colour = "black"))

```


